```{r initiate-environment-v1, echo=TRUE, warning=FALSE}
### Load visit 1 analysis specific packages
# Set up required packages
.cran_packages <- c("ggpubr", "mclust", "dunn.test", "cluster")
.bioc_packages <- c()

.inst <- .cran_packages %in% installed.packages()
if(any(!.inst)) {
  install.packages(.cran_packages[!.inst])
}
.inst <- .bioc_packages %in% installed.packages()
if(any(!.inst)) {
  source("http://bioconductor.org/biocLite.R")
  biocLite(.bioc_packages[!.inst], ask = F)
}
# Load packages into session, and print package version
sapply(c(.cran_packages, .bioc_packages), require, character.only = TRUE)
```

### Sample sequencing depth

```{r visit-one-data}
## Sequencing depth summary
dt.ps2.v1v2v3.sample_data = cbind(as(sample_data(ps2.v1v2v3), "data.frame"),
                          TotalReads = sample_sums(ps2.v1v2v3)) ##Total reads is sequence depth
gg.SeqDepth = ggplot(dt.ps2.v1v2v3.sample_data, aes(TotalReads)) + geom_histogram() + ggtitle("Sequencing Depth")
ggplotly(gg.SeqDepth)

```
The graph shows how many reads each sample has. The Y-axis shows the number samples and the X-axis shows the total number of reads for those samples

### Sample distribution accross visits

How are the participants distributed accross visits?

```{r participants-dist-accross-visit}
## Distribution of BV accross visits
cat("BV status")
table(sample_data(ps2.v1v2v3)$bvscat, sample_data(ps2.v1v2v3)$VisitCode)


cat("\nInflammation")
table(sample_data(ps2.v1v2v3)$Inflammation, sample_data(ps2.v1v2v3)$VisitCode)

cat("\nInflammation vs BV")
table(sample_data(ps2.v1v2v3)$Inflammation, sample_data(ps2.v1v2v3)$bvscat)

cat("\nSTI")
table(sample_data(ps2.v1v2v3)$STI, sample_data(ps2.v1v2v3)$VisitCode)
```

## Estimate richness and transform to relative abundance

```{r }
## Estimate richness
ps2.v1v2v3.rich <- cestimate_richness(ps2.v1v2v3)

## Aglomerate to species
ps2.v1v2v3.glom.species <- tax_glom(ps2.v1v2v3.rich, "Species")

## Transform to relative abundance
ps2.v1v2v3.glom.species.ra <- transform_sample_counts(ps2.v1v2v3.glom.species, function(x) {
  x/sum(x)
})
otu_table(ps2.v1v2v3.glom.species.ra)[is.nan(otu_table(ps2.v1v2v3.glom.species.ra))] <- 0

## Distribution of samples accross CSTS
levels(sample_data(ps2.v1v2v3.glom.species.ra)$sim_CST)

table(sample_data(ps2.v1v2v3.glom.species.ra)$sim_CST)

table(sample_data(ps2.v1v2v3.glom.species.ra)$sim_CST, sample_data(ps2.v1v2v3.glom.species.ra)$VisitCode)

table(sample_data(ps2.v1v2v3.glom.species.ra)$sim_CST, sample_data(ps2.v1v2v3.glom.species.ra)$bvscat)

#Sort by CTS snd Shannon indexes before proceeding
sample_data(ps2.v1v2v3.glom.species.ra) <- sample_data(ps2.v1v2v3.glom.species.ra)[order(sample_data(ps2.v1v2v3.glom.species.ra)$sim_CST, sample_data(ps2.v1v2v3.glom.species.ra)$ShannonLgT), ]
```

## Alpha and beta diversity ordinations

```{r echo=FALSE}

# PCoA plots -color: bvscat

ps2.v1v2v3.glom.species.ra <- prune_samples(sample_sums(ps2.v1v2v3.glom.species.ra) > 0, ps2.v1v2v3.glom.species.ra)
ps2.v1v2v3.glom.species.ra <- prune_taxa(taxa_sums(ps2.v1v2v3.glom.species.ra) > 0, ps2.v1v2v3.glom.species.ra)

ord.ps2.v1v2v3.glom.species.ra.pcoa.bray <- ordinate(ps2.v1v2v3.glom.species.ra, method = "PCoA", distance = "bray")
p.ord.ps2.v1v2v3.glom.species.ra.pcoa.bray <- plot_ordination(ps2.v1v2v3.glom.species.ra, 
                                                              ord.ps2.v1v2v3.glom.species.ra.pcoa.bray, color = "SeqRun", shape = "STI", type = "samples", 
                                                              title = "Bray Curtis Accross Visits") + 
  theme_light()
p.ord.ps2.v1v2v3.glom.species.ra.pcoa.bray + facet_wrap(~VisitCode) + geom_point(size=1)

# PCoA plots -color: visitcode
sample_data(ps2.v1v2v3.glom.species.ra)$RowNo <- paste(seq_along(1 : nrow(sample_data(ps2.v1v2v3.glom.species.ra))))
p.ord.ps2.v1v2v3.glom.species.ra.pcoa.bray.visits <- plot_ordination(ps2.v1v2v3.glom.species.ra, 
                                                                     ord.ps2.v1v2v3.glom.species.ra.pcoa.bray, color = "bvscat", shape = "STI", 
                                                                     type = "samples", title = "Bray Curtis Accross Visits") +
  #geom_text_repel(aes(label = RowNo), direction = "x") +
  theme_light() 
p.ord.ps2.v1v2v3.glom.species.ra.pcoa.bray.visits + geom_point(size=1)

## Create factor levels for visits
sample_data(ps2.v1v2v3.glom.species.ra)$VisitCode <- factor(sample_data(ps2.v1v2v3.glom.species.ra)$VisitCode, levels = c(1000, 1020, 1030))

# Create factor variable for BV clearence
table(sample_data(ps2.v1v2v3.glom.species.ra)$bvscat, sample_data(ps2.v1v2v3.glom.species.ra)$VisitCode)
bv.negative.at.start <- sample_data(ps2.v1v2v3.glom.species.ra)$ParticipantID[sample_data(ps2.v1v2v3.glom.species.ra)$VisitCode == 1000 & sample_data(ps2.v1v2v3.glom.species.ra)$bvscat == "Negative"]
bv.negative.at.start

```
### Change in Shannon Diversity accross visits (Overall significance)

```{r}

sd.ps2.v1v2v3.glom.species.ra <- data.frame(sample_data(ps2.v1v2v3.glom.species.ra))
ggpaired(sd.ps2.v1v2v3.glom.species.ra, x = "VisitCode", y = "ShannonLgT",
         color = "VisitCode", line.color = "gray", line.size = 0.4,
         palette = visitsPallet, id = "ParticipantID", width = 0.1 ) +
  scale_x_discrete(breaks=c(1000, 1020, 1030), labels=c("Baseline", "Visit 2", "Visit 3")) +
  scale_color_discrete(name = "Clinical Visits", labels = c("Baseline", "Visit 2", "Visit 3")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_light() +
  ylab("Shannon Diversity") +
  xlab("Clinical Visits") +
  stat_compare_means(paired = TRUE)

```

### Change in Shannon diversity levels accross visits (compare visits)

```{r}

## Change in Shannon diversity accross visits
my_comparisons <- list( c("1000", "1020"), c("1020", "1030"))
ggpaired(sd.ps2.v1v2v3.glom.species.ra, x = "VisitCode", y = "ShannonLgT",
         color = "VisitCode", line.color = "gray", line.size = 0.4,
         palette = visitsPallet, id = "ParticipantID", width = 0.1 ) +
  scale_x_discrete(breaks=c(1000, 1020, 1030), labels=c("Baseline", "Visit 2", "Visit 3")) +
  scale_color_discrete(name = "Clinical Visits", labels = c("Baseline", "Visit 2", "Visit 3")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_light() +
  ylab("Shannon Diversity") +
  xlab("Clinical Visits") +
  stat_compare_means(paired = TRUE, comparisons = my_comparisons)

```

PS: Legends for the annotations will be imposed in the final graphics using a different program. However, for BV, No BV=white, intermediate=grey and bv positive=maroon. For the rest of the annotations, color=positive.

```{r alpha-diverstiy-plots, echo=FALSE, ,  fig.width= 8}


#Plot log10 transformed shannon indexes
bp <- ggplot(sd.ps2.v1v2v3.glom.species.ra, aes(x = (factor(sd.ps2.v1v2v3.glom.species.ra$SampleID, levels = sample.order)), y = ShannonLgT, color = CSTs)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(CSTsPallete) +
  xlab("Samples") +
  theme_classic() +
  theme(axis.text.x=element_blank())

# Observed
gg.ps2.v1v2v3.observed <- ggplot(sd.ps2.v1v2v3.glom.species.ra, aes(x = STI, y = Observed,  color = bvscat)) +
  stat_smooth(method = "loess") +
  geom_boxplot() +
  labs(y = "Observed", color = "bvscat", x = "STI") +
  #geom_jitter(size = 2, alpha = 0.5, width = 0.5) +
  theme(legend.position="bottom")# +# +
  #scale_color_manual(values = c("black", "chocolate", "green", "purple", "blue"))

# Shannon diversity
gg.ps2.v1v2v3.shannon <- ggplot(sd.ps2.v1v2v3.glom.species.ra, aes(x = STI, y = ShannonLgT, color = bvscat)) +
  stat_smooth(method = "loess") +
  geom_boxplot(alpha=0.1) +
  labs(y = "Shannon", color = "Any STI", x = "STI") +
  #geom_jitter(size = 2, alpha = 0.5, width = 0.5) +
  theme(legend.position="bottom")# + #+
  #scale_color_manual(values = c("black", "chocolate", "green", "purple", "blue"))

# Simpson diversity
gg.ps2.v1v2v3.simpson <- ggplot(sd.ps2.v1v2v3.glom.species.ra, aes(x = STI, y = Simpson, color = bvscat)) +
  #stat_smooth(method = "loess") +
  geom_boxplot( alpha=0.1) +
  labs(y = "Simpson", color = "Any STI", x = "STI") +
  #geom_jitter(size = 2, alpha = 0.5, width = 0.5) +
  theme(legend.position="bottom")# +
  #scale_color_manual(values = c("black", "chocolate", "green", "purple", "blue"))

# Note: ggarrange enables you to directly add labels to plot grids
grid.arrange(gg.ps2.v1v2v3.observed, gg.ps2.v1v2v3.shannon, gg.ps2.v1v2v3.simpson, ncol = 1, nrow = 3)

#A quick test for normalcy
shapiro.test(sd.ps2.v1v2v3.glom.species.ra$ShannonLgT)
hist(sd.ps2.v1v2v3.glom.species.ra$ShannonLgT, main = "Histogram of Log10 Transformed Shannon Diversity Indexes", xlab = "Shannon (Log10 Transformed)")

## Statistical analysis on alpha diversity

#By BVCat
kruskal.test(sd.ps2.v1v2v3.glom.species.ra$ShannonLgT , as.factor(sd.ps2.v1v2v3.glom.species.ra$bvscat))
dunn.test(sd.ps2.v1v2v3.glom.species.ra$ShannonLgT, sd.ps2.v1v2v3.glom.species.ra$bvscat)
ggplot(sd.ps2.v1v2v3.glom.species.ra, aes(bvscat, ShannonLgT, color = bvscat)) +
  geom_boxplot() +
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  facet_grid("VisitCode") +
  ylab("Shannon (Log10 transformed)") +
  xlab("BV Category")

#By CST
kruskal.test(sd.ps2.v1v2v3.glom.species.ra$ShannonLgT, sd.ps2.v1v2v3.glom.species.ra$sim_CST)
dunn.test(sd.ps2.v1v2v3.glom.species.ra$ShannonLgT, sd.ps2.v1v2v3.glom.species.ra$sim_CST)
ggplot(sd.ps2.v1v2v3.glom.species.ra, aes(sim_CST, ShannonLgT, label = Position)) +
  geom_boxplot() +
  geom_jitter(shape=16, position=position_jitter(0.2), aes(color = SeqRun))  +
  facet_grid("VisitCode") +
  ylab("Shannon (Log10 transformed)") +
  xlab("CST")

#By STI
wilcox.test(ShannonLgT ~ STI, data = sd.ps2.v1v2v3.glom.species.ra)
wilcox.test(ShannonLgT ~ Inflammation, data = sd.ps2.v1v2v3.glom.species.ra)

#By PSA
wilcox.test(ShannonLgT ~ PSA, data = sd.ps2.v1v2v3.glom.species.ra)


#Mixed models to see combined effects
glm(formula = ShannonLgT ~ bvscat * Inflammation, data = sd.ps2.v1v2v3.glom.species.ra)
```

## Betadiversity

### Batch effects - sequencing run

Do we see any effect of sequencing run that can confound our analysis?

```{r}
evals <- ord.ps2.v1v2v3.glom.species.ra.pcoa.bray$values$Eigenvalues
p.pcoa.bray <- plot_ordination(ps2.v1v2v3.glom.species.ra, ord.ps2.v1v2v3.glom.species.ra.pcoa.bray, color = "SeqRun", axes = c(1,2)) +
  geom_point(size = 2) +
  scale_fill_manual(values = CSTsPallete) +
  labs(title = "PCoA of Bray Curtis (Abundance)", color = "SeqRuns") +
  theme_minimal() +
  coord_fixed(sqrt(evals[2] / evals[1])) #+
  #stat_ellipse(type = "t")
p.pcoa.bray
```

### Do we see any clustering CSTs

```{r plot-ordination, echo=FALSE}

# Plot the data by the treatment variable
y = 1000 # Set a threshold for the minimum number of acceptable reads. Can start as a guess
x = "bvscat" # Set the x-axis variable you want to examine
label = "SampleID" # This is the label you want to overlay on the points

ord.pcoa.jac <- ordinate(ps2.v1v2v3.glom.species.ra, method = "PCoA", distance = "jaccard")
p.pcoa.jac <- plot_ordination(ps2.v1v2v3.glom.species.ra, ord.pcoa.jac, color = "sim_CST" , axes = c(1,2)) +
  geom_point(size = 1) +
  scale_fill_manual("CSTs", breaks = CSTs_all, values = CSTsPalleteNamed) +
  labs(title = "PCoA of Jacard (Presence/Absence)", color = "CSTs") +
  stat_ellipse(type = "t") +
  theme_minimal()

p.pcoa.bray <- plot_ordination(ps2.v1v2v3.glom.species.ra, ord.ps2.v1v2v3.glom.species.ra.pcoa.bray, color = "sim_CST", axes = c(1,2)) +
  geom_point(size = 1) +
  scale_fill_manual("CSTs", breaks = CSTs_all, values = CSTsPalleteNamed) +
  labs(title = "PCoA of Bray Curtis (Abundance)", color = "CSTs") +
  stat_ellipse(type = "t") +
  theme_minimal()

ord.pcoa.unif <- ordinate(ps2.v1v2v3.glom.species.ra, method = "PCoA", distance = "unifrac")
p.pcoa.unif <- plot_ordination(ps2.v1v2v3.glom.species.ra, ord.pcoa.unif, color = "sim_CST" , axes = c(1,2)) +
  scale_fill_manual("CSTs", breaks = CSTs_all, values = CSTsPalleteNamed) +
  geom_point(size = 1) +
  labs(title = "PCoA of Unifrac Distance", color = "CSTs") +
  stat_ellipse(type = "t") +
  theme_minimal()

ord.pcoa.wunif <- ordinate(ps2.v1v2v3.glom.species.ra, method = "PCoA", distance = "wunifrac")
p.pcoa.wunif <- plot_ordination(ps2.v1v2v3.glom.species.ra, ord.pcoa.wunif, color = "sim_CST", axes = c(1,2)) +
  scale_fill_manual("CSTs", breaks = CSTs_all, values = CSTsPalleteNamed) +
  geom_point(size = 1) +
  labs(title = "PCoA of W. Unifrac", color = "CSTs") +
  stat_ellipse(type = "t") +
  theme_minimal()


grid.arrange(p.pcoa.jac, p.pcoa.bray, p.pcoa.unif, p.pcoa.wunif, ncol = 2, nrow = 2)

cat("Cross tabulation of CTs and BV Score")
table(sample_data(ps2.v1v2v3.glom.species.ra)$sim_CST, sample_data(ps2.v1v2v3.glom.species.ra)$bvscat)

cat("Cross tabulation of CTs and Inflamation")
table(sample_data(ps2.v1v2v3.glom.species.ra)$sim_CST, sample_data(ps2.v1v2v3.glom.species.ra)$Inflammation)
```
